Paper figures

### 2019-01-09 paper_figures.Rmd
Changed fig dimension

```{r supplfig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3'}
filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'Chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value, fill=timepoint)) +
  scale_fill_manual(name = "Timepoint", values = cm_palette) +
  xlab('Timepoint') +
  ylab('Value') +
  facet_grid(variable~.,scales = 'free_y') +
  guides(fill=FALSE) +
  theme_cm()
```

@@

```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3"}
subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
  theme_cm() +
  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))),
    gp=grid::gpar(col="black", fontsize=13)),  xmin = 0.15, xmax = .15, ymin = .05, ymax = .05)
```

@@

```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families."}
  lacno_bact_relabs %>%
  ggplot() +
    geom_boxplot(aes(timepoint, Abundance, fill=timepoint)) + 
    scale_fill_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
    facet_wrap(interaction(Family, Genus) ~ ., scales = "free_x", as.table= T) +
    guides(fill=FALSE) +
    theme_cm()
```

```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints"}
not12 <- gg_topi %>%
  subset_samples(!(timepoint %in% c("T1", "T2")))

@@

## Panel ?: Ordination beta diversity T3 vs T7
```{r}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

@@

```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered"}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(OTU %in% (genus_ab_t %>%
  group_by(OTU, timepoint) %>%
  summarise(median=median(Abundance)) %>%
  group_by(OTU) %>%
  summarise(filt=min(median)>0.01) %>% filter(filt==TRUE))$OTU) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85)) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm()
```

```{r supplfig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis"}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85)) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  scale_y_log10() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm() +
  theme(strip.text.y = element_text(angle = 360),
        legend.position = "bottom")
```

@@

otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
ggplot() +
  geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
  scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
  ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 0.5, tip_length = 0, test = 't.test') +
  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free", shrink = FALSE) +
  theme_cm() +
  ylab('Relative abundance') +
  xlab('Timepoint')
```

@@

```{r fig3G, corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  mutate(timepoint = paste('T',timepoint2,'-T3',sep = '')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() +
  facet_wrap(facets = 'timepoint', nrow = 1, shrink = TRUE, scales = 'free_y') +
  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
  ylab(bquote('Tumor mass'~( mm^3))) +
  xlab('Bray Curtis distance') +
  theme_cm() +
  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label), inherit.aes = FALSE)
```

```{r corr tumor braycurtis, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity "}

otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  filter(timepoint %in% c('T3','T7')) %>%
  select(-SampleID, -OTU) %>%
  group_by(condition) %>%
  summarise(Abundance = list(Abundance)) %>%
  spread(condition, Abundance)

  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus)))

>>>>

```{r supplfig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.dim = c(5,5)}
filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Observed', 'Chao1', 'Simpson', 'Shannon')) %>%
  ggplot() +
  geom_boxplot(aes(timepoint, value, fill=timepoint)) +
  scale_fill_manual(name = "Timepoint", values = cm_palette) +
  xlab('Timepoint') +
  ylab('Value') +
  facet_grid(variable~.,scales = 'free_y') +
  guides(fill=FALSE) +
  theme_cm()
```

@@

```{r Plot ordination, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(5,5)}
subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
  theme_cm() +
  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))),
    gp=grid::gpar(col="black", fontsize=11)),  xmin = 0.15, xmax = .15, ymin = .12, ymax = .12)
```

@@

```{r fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(4,4)}
  lacno_bact_relabs %>%
  ggplot() +
    geom_boxplot(aes(timepoint, Abundance, fill=timepoint)) + 
    scale_fill_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
    facet_wrap(interaction(Family, Genus) ~ ., scales = "free_x", as.table= T) +
    guides(fill=FALSE) +
    theme_cm()
```

```{r fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(5,6)}
not12 <- gg_topi %>%
  subset_samples(!(timepoint %in% c("T1", "T2")))

@@

## Panel ?: Ordination beta diversity T3 vs T7
```{r, fig.dim = c(5, 6)}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

@@

```{r fig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=8}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(OTU %in% (genus_ab_t %>%
  group_by(OTU, timepoint) %>%
  summarise(median=median(Abundance)) %>%
  group_by(OTU) %>%
  summarise(filt=min(median)>0.01) %>% filter(filt==TRUE))$OTU) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85)) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm()
```

```{r supplfig3E, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85)) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  scale_y_log10() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm() +
  theme(strip.text.y = element_text(angle = 360),
        legend.position = "bottom")
```

@@

otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
ggplot() +
  geom_boxplot(aes(timepoint, Abundance, fill=condition)) +
  scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free", shrink = FALSE) +
  theme_cm() +
  ylab('Relative abundance') +
  xlab('Timepoint')
```

@@

```{r fig3G, corr tumor braycurtis, fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  mutate(timepoint = paste('T',timepoint2,'-T3',sep = '')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() +
  facet_wrap(facets = 'timepoint', nrow = 1, shrink = TRUE, scales = 'free_y') +
  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
  ylab(bquote('Tumor mass'~( mm^3))) +
  xlab('Bray Curtis distance') +
  theme_cm() +
  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label), inherit.aes = FALSE)
```

```{r corr tumor braycurtis, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
tbl_data <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tidyr::gather_("SampleID", "Abundance", colnames(.)) %>%
  mutate(mouseid=stringr::str_sub(.$SampleID,8, 10)) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  filter(timepoint %in% c('T3','T7')) %>%
  select(-SampleID) %>%
  spread(timepoint,Abundance) %>%
  mutate(diff_abb = T3-T7) %>%
  arrange(mouseid) %>%
  bind_cols( gg_topi.genus.relab %>%
              sample_data %>%
              filter(timepoint == 'T7') %>%
              select(id.topo, tumor_mass) %>%
              arrange(id.topo) %>%
              select(-id.topo))

tbl_data %>%
ggplot(aes(diff_abb, tumor_mass)) +
  geom_point() +
  facet_wrap(facets = 'condition', nrow = 1, scales = 'free') +
  stat_smooth(method = 'lm', color='red') + 
  theme_cm() +
  xlab('Relative abundance (T3-T7)') +
  ylab(bquote('Tumor mass'~( mm^3)))
```

```{r}
tbl_data %>%
  select(-T3,-T7, -mouseid) %>%
  group_by(condition) %>%
  summarise(diff_abb = list(diff_abb), tumor_mass = list(tumor_mass)) %>%
  group_by(condition) %>%
  mutate(cor.test = paste(unlist(cor.test(unlist(diff_abb), unlist(tumor_mass))[c('p.value','estimate')]), collapse = ' ')) %>%
  separate(cor.test, into = c('p.value','estimate'), sep = ' ')
```
